Fail Faster

### 2025-03-24 paper/paper.tex
blah

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Property-based testing relies on random input generators to uncover bugs, often under tight time constraints. These generators are built 
  using generator libraries, which provide expressive combinators for constructing complex inputs. 
  Although convenient,
  today's generator libraries suffer from substantial performance costs, limited by abstraction overhead and inefficient randomness libraries.
  In this paper, we identify and address these bottlenecks. We introduce \name, a multi-stage programming technique that eliminates generator overhead
  without sacrificing the flexibility and expressiveness of existing generator library functions. We augment leading generator libraries in OCaml and 
  Scala 3 to use the \name approach, preserving the original library's semantics and enabling precise, pointwise performance comparisons. We also show that \rand
  has a measurable impact on generator performance, quantify its contribution, and show that speedups of $5.5\times$ are possible. Together, these 
  improvements lead to bug-finding up to $13\times$ faster---showing that high-level generator abstractions need not come at the cost of speed.
\end{abstract}

@@

This is due to two major sources of inefficiency.
First, the same high-level design approach that gives generator libraries their
flexibility also introduces layered abstractions and indirections that are difficult for
compilers to optimize.
Second, the calls a generator makes to its \rand{} can constitute a large 
proportion of its running time, magnifying the cost of an inefficient 
implementation.

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Property-based testing relies on random input generators to uncover bugs, often under tight time constraints. These generators are built 
  using generator libraries, which provide expressive combinators for constructing complex inputs. 
  Although convenient,
  today's generator libraries suffer from substantial performance costs, limited by abstraction overhead and inefficient randomness libraries.
  In this paper, we identify and address these bottlenecks. We introduce \name, a multi-stage programming technique that eliminates generator overhead
  without sacrificing the flexibility and expressiveness of existing generator library functions. We augment leading generator libraries in OCaml and 
  Scala 3 to use the \name approach, preserving the original library's semantics and enabling precise, pointwise performance comparisons. We also show that \rand
  has a measurable impact on generator performance, quantify its contribution, and show that speedups of $5.5\times$ are possible. Together, these 
  improvements lead to bug-finding up to $13\times$ faster---showing that high-level generator abstractions need not come at the cost of speed. \tr{Needs a lot of work. No spice.}
\end{abstract}

@@

This is due to two major sources of inefficiency.
First, the same high-level design approach that gives generator libraries their
flexibility also introduces layered abstractions and indirections that are difficult for
compilers to optimize.
Second, the calls a generator makes to its \rand{} can constitute a large 
proportion of its run-time, magnifying the cost of an inefficient 
implementation.
